Empirical Evaluation of eval in R

### 2020-08-19 inst/analysis/analysis_stable.Rmd
Classification of calls on expr_resolved

---
title: "Empirical Evaluation of eval in R"
output:
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
    data_filepath: "/var/lib/R/project-evalR/run/package-evals-traced-corpus.2/calls.fst"
    #data_filepath: "/var/lib/R/project-evalR/run/package-evals-traced-all.1/calls.fst"
    # The trace data on all CRAN does not have all the columns of the smaller corpus. The analysis will fail for the big corpus.

@@

- assignments
- reading (in a data frame, a vector)
- parsing
- defusing (`quote` and so on)
- conversion or creation of new values
- building calls and calling
- working with environments
- class assignments, reads and definitions
- building logical values
- formula
- meta programming on functions (getting access to the body of the function, to the call stack)
- 'NA' corresponds to no function calls (or that it was not parsed successfully)

```{r group_expr_functions}
group_expr_functions <- expr_call_arg %>% mutate(group_function_expr = fct_collapse(function_expr, 
                          assignment = c("<-", "[[<-", "[<-", "$<-"),
                          read = c("[", "[[", "$"),
                          parsing = c("parse", "str2lang", "str2expression", "parse_only", "base::parse"),
                          defusing = c("quote", "substitute", "bquote", "expr", "quo"),
                          expressions = c("as.list", "expression", "as.name", "as.expression", "list", "as.data.frame", "as.character", "{", "modify_lang"),
                          calling = c("call", "as.call", "do.call"),
                          environment = c("environment", "as.environment"),
                          classAssign = c("@<-"),
                          classRead = c("slot", "@"),
                          classDef = c(".Primitive(\"class\")"),
                          logical = c("<", ">", "is.element"),
                          choice = c("if"),
                          function_meta = c("sys.calls", "function", "body"),
                          formula = c("~", "deriv", "D", "model.frame", "rootftfunc", "contest.lmerModLmerTest", "JacFunc", "gamlss.family", "contest1D.lmerModLmerTest", "contestMD.lmerModLmerTest", "reexpr"),                           
                          other_level = "other" ))
                          #TODO: add anonymous function applications
```

@@

>>>>

---
title: "Empirical Evaluation of eval in R"
output:
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
    data_filepath: "/var/lib/R/project-evalR/run/package-evals-traced-corpus.3/calls.fst"
    #data_filepath: "/var/lib/R/project-evalR/run/package-evals-traced-all.1/calls.fst"
    # The trace data on all CRAN does not have all the columns of the smaller corpus. The analysis will fail for the big corpus.

@@

- assignments
- reading (in a data frame, a vector)
- parsing
- defusing (`quote` and so on)
- conversion or creation of new values
- building calls and calling
- working with environments
- class assignments, reads and definitions
- building logical values
- formula
- meta programming on functions (getting access to the body of the function, to the call stack)
- 'NA' corresponds to no function calls (or that it was not parsed successfully)
- anonymous functions

```{r group_expr_functions}
group_expr_functions <- expr_call_arg %>%
                            mutate(function_expr2 = map_chr(detect_anonymous_func, function_expr)) %>%
                            mutate(group_function_expr = fct_collapse(function_expr2, 
                              assignment = c("<-", "[[<-", "[<-", "$<-"),
                              read = c("[", "[[", "$"),
                              parsing = c("parse", "str2lang", "str2expression", "parse_only", "base::parse"),
                              defusing = c("quote", "substitute", "bquote", "expr", "quo"),
                              expressions = c("as.list", "expression", "as.name", "as.expression", "list", "as.data.frame", "as.character", "{", "modify_lang"),
                              calling = c("call", "as.call", "do.call"),
                              environment = c("environment", "as.environment"),
                              classAssign = c("@<-"),
                              classRead = c("slot", "@"),
                              classDef = c(".Primitive(\"class\")"),
                              logical = c("<", ">", "is.element"),
                              choice = c("if"),
                              anonymous = c("anonymous"),
                              function_meta = c("sys.calls", "function", "body"),
                              formula = c("~", "deriv", "D", "model.frame", "rootftfunc", "contest.lmerModLmerTest", "JacFunc", "gamlss.family", "contest1D.lmerModLmerTest", "contestMD.lmerModLmerTest", "reexpr"),                           
                              other_level = "other" ))
```

@@

We can do the same analysis as for the non-resolved expressions:

```{r expr_resolved__call_arg}
expr_resolved_call_arg <- eval_calls %>% filter(expr_resolved_type == "LANGSXP") %>% 
  mutate(function_expr = map_chr(expr_resolved, function_name))
```

```{r top_level_function_resolved_expr}
expr_resolved_call_arg %>% count(function_expr) %>% arrange(desc(n))
```

```{r group_expr_resolved_functions}
group_expr_resolved_functions <- expr_resolved_call_arg %>%
                            mutate(function_expr2 = map_chr(detect_anonymous_func, function_expr)) %>%
                            mutate(group_function_expr = fct_collapse(function_expr2, 
                              assignment = c("<-", "[[<-", "[<-", "$<-"),
                              read = c("[", "[[", "$"),
                              parsing = c("parse", "str2lang", "str2expression", "parse_only", "base::parse"),
                              defusing = c("quote", "substitute", "bquote", "expr", "quo"),
                              expressions = c("as.list", "expression", "as.name", "as.expression", "list", "as.data.frame", "as.character", "{", "modify_lang"),
                              calling = c("call", "as.call", "do.call"),
                              environment = c("environment", "as.environment"),
                              classAssign = c("@<-"),
                              classRead = c("slot", "@"),
                              classDef = c(".Primitive(\"class\")"),
                              logical = c("<", ">", "is.element"),
                              choice = c("if"),
                              anonymous = c("anonymous"),
                              function_meta = c("sys.calls", "function", "body"),
                              formula = c("~", "deriv", "D", "model.frame", "rootftfunc", "contest.lmerModLmerTest", "JacFunc", "gamlss.family", "contest1D.lmerModLmerTest", "contestMD.lmerModLmerTest", "reexpr"),                           
                              other_level = "other" ))
```

```{r}
group_expr_resolved_functions %>% count(group_function_expr) %>% arrange(desc(n)) %>% knitr::kable()
```

```{r}
group_expr_resolved_functions %>% ggplot() +
  geom_bar(aes(x = fct_rev(fct_infreq(group_function_expr)))) +
  scale_y_sqrt() +
  coord_flip() +
  labs(title = "Category of calls", y = "n", x = "category" ) 
```
